Inferring Galactic Parameters from Chemical Abundances with Simulation-Based Inference

### 2025-02-17 src/tex/ms.tex
[showyourwork] overleaf sync

Our mock data generation and inference then follows the one of Sec.~\ref{subsec:chempy_tng}. This means we apply our NPE trained on data with the TNG yield set to data created with the above mentioned alternative yield set. This effectively probes the effect of model missspecification on the inference results.

We show our inference results for this setup for varying number of stars (blue) in Fig.~\ref{fig:CHEMPY_alt_N_star_analysis} comparing again against HMC results (red) from \citet{Philcox_2019}. We see that the SBI results are less biased than the HMC results especially for $\alpha_{\rm IMF}$. In fact with 200 stars HMC inferences are about \textcolor{red}{calculate value} $\sigma$ away from the ground truth value for $\alpha_{\rm IMF}$ and $\log_{10}(N_{\rm Ia})$, respectively. Again, the joint posterior is shown in Fig.~\ref{fig:CHEMPY_alt_sbi} and shows that while individually parameter inferences look good, jointly taken the SBI results are on the edge of being $3\sigma$ away from the ground truth.

>>>>

Our mock data generation and inference then follows the one of Sec.~\ref{subsec:chempy_tng}. This means we apply our NPE that we trained on \texttt{CHEMPY} stellar abundances simulated with the Ilustris TNG yield set to a set of stellar abundances simulated with \texttt{CHEMPY} but using the above mentioned alternative yield set. This effectively probes the effect of model missspecification on the inference results.

We show our inference results for this setup for varying number of stars (blue) in Fig.~\ref{fig:CHEMPY_alt_N_star_analysis} comparing again against HMC results (red) from \citet{Philcox_2019}. We see that the SBI results are less biased than the HMC results especially for $\alpha_{\rm IMF}$. In fact with 100 stars HMC inferences are about \textcolor{red}{calculate value} $\sigma$ away from the ground truth value for $\alpha_{\rm IMF}$ and $\log_{10}(N_{\rm Ia})$, respectively. Again, the joint posterior is shown in Fig.~\ref{fig:CHEMPY_alt_sbi} and shows that while individually parameter inferences look good, jointly taken the SBI results are on the edge of being $3\sigma$ away from the ground truth.
